---
title: "R Workshop Report"
author: "Cheng Yew Toon, Zaw Myo Tun, Maria Baig"
format: html
date: today
date-format: "DD MMMM YYYY"
execute: 
  echo: false
  warning: false
  message: false
---

# Report for R Workshop (January 22 to 26, 2024)

```{r}
#| message : false
#| warning: false

# Plots for Likert chart

# Load package
library(tidyverse)
library(googlesheets4)
library(collapse)

# Set working directory
here::set_here("/Users/cheng/OneDrive - ird.global/Desktop/R Projects/R Workshop Feedback/R_workshop_feedback")

# Load data
url <- "https://docs.google.com/spreadsheets/d/12jxEV9rqIOrIw70lVXfSegorv25owciTMS-hjSotioo/edit?resourcekey#gid=1120526525"
dat <- read_sheet(url)
r_session <- read_csv("R Workshop January 2024 (participant characteristics) - Copy of Attended for 1 day.csv")

dat1 <- dat |> 
  rename(email = 2) |> 
  filter(row_number() == n(), .by = email) |> # Take the last record if duplicated
  select(-Timestamp)

respondents_count <- nrow(dat1)

#load data for pre-workshop organisation
pre_wksp <- dat1 |> 
  select(email, matches("1.[1-4]")) |> 
  pivot_longer(-email,
               names_to = "qns",
               values_to = "rating") |> 
  mutate(
    qns = str_sub(qns, 3, 3) |> as.numeric(), # 3 is the index position of the str_sub for qns
    qns = case_when(
      qns == 1 ~ "It was easy to register for the workshop",
      qns == 2 ~ "Communications for the workshop were clear",
      qns == 3 ~ "Communications for the workshop were timely",
      qns == 4 ~ "Instructions for workshop preparation were clear",
      .default = "Unknown"
    ),
    qns = factor(qns,
                 levels = c("It was easy to register for the workshop","Communications for the workshop were clear",
                            "Communications for the workshop were timely","Instructions for workshop preparation were clear")),
    rating = case_when(
      rating == 1 ~ "Strongly disagree",
      rating == 2 ~ "Disagree",
      rating == 3 ~ "Neutral",
      rating == 4 ~ "Agree",
      rating == 5 ~ "Strongly agree",
      .default = "Unknown"
    ),
    rating = factor(rating,
                    levels = c("Strongly disagree", "Disagree", "Neutral",
                               "Agree", "Strongly agree"))
  )

pre_wksp_1 <- pre_wksp |> 
  count(qns, rating) |> 
  mutate(prop = n/sum(n),
         .by = qns)

# load data for Facilitator explanation
explain <- dat1 |> 
  select(email, matches("^[2-7]\\.1")) |> 
  pivot_longer(-email,
               names_to = "topic",
               values_to = "rating") |> 
  mutate(
    topic = str_sub(topic, 1, 1) |> as.numeric(),
    topic = case_when(
      topic == 2 ~ "Intro to R",
      topic == 3 ~ "Data Viz",
      topic == 4 ~ "Data Wrangling",
      topic == 5 ~ "Descriptive Analysis",
      topic == 6 ~ "Intro to Quarto",
      topic == 7 ~ "Report generation",
      .default = "Unknown"
    ),
    topic = factor(topic,
                   levels = c("Intro to R", "Data Viz", "Data Wrangling",
                              "Descriptive Analysis", "Intro to Quarto",
                              "Report generation")),
    rating = case_when(
      rating == 1 ~ "Strongly disagree",
      rating == 2 ~ "Disagree",
      rating == 3 ~ "Neutral",
      rating == 4 ~ "Agree",
      rating == 5 ~ "Strongly agree",
      .default = "Unknown"
    ),
    rating = factor(rating,
                    levels = c("Strongly disagree", "Disagree", "Neutral",
                               "Agree", "Strongly agree")),
    qns = "The facilitator's explanation was easy to follow"
  )

explain_1 <- explain |> 
  count(topic, rating, qns) |> 
  mutate(prop = n/sum(n),
         .by = topic)

# load data for materials
materials <- dat1 |> 
  select(email, matches("^[2-7]\\.2")) |> 
  pivot_longer(-email,
               names_to = "topic",
               values_to = "rating") |> 
  mutate(
    topic = str_sub(topic, 1, 1) |> as.numeric(),
    topic = case_when(
      topic == 2 ~ "Intro to R",
      topic == 3 ~ "Data Viz",
      topic == 4 ~ "Data Wrangling",
      topic == 5 ~ "Descriptive Analysis",
      topic == 6 ~ "Intro to Quarto",
      topic == 7 ~ "Report generation",
      .default = "Unknown"
    ),
    topic = factor(topic,
                   levels = c("Intro to R", "Data Viz", "Data Wrangling",
                              "Descriptive Analysis", "Intro to Quarto",
                              "Report generation")),
    rating = case_when(
      rating == 1 ~ "Strongly disagree",
      rating == 2 ~ "Disagree",
      rating == 3 ~ "Neutral",
      rating == 4 ~ "Agree",
      rating == 5 ~ "Strongly agree",
      .default = "Unknown"
    ),
    rating = factor(rating,
                    levels = c("Strongly disagree", "Disagree", "Neutral",
                               "Agree", "Strongly agree")),
    qns = "The training materials clear and adequate"
  )

materials_1 <- materials |> 
  count(topic, rating, qns) |> 
  mutate(prop = n/sum(n),
         .by = topic)

p_pre_wksp <-  ggplot(pre_wksp_1, aes(x = prop, y = fct_rev(qns), fill = fct_rev(rating))) +
  geom_col(alpha = 0.8, show.legend = TRUE) +
  labs(x = "", y = "", fill = "",
       title = "Pre-workshop Organisation") +
  scale_fill_manual(values = c("#236EC3", "#78AFE6", "#E1E1E1", "#E6AAA0", "#C32314")) +
  guides(fill = guide_legend(reverse = TRUE)) + # reverse legend order
  scale_x_continuous(labels = scales::label_percent()) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank()   
  )

p_explain <- ggplot(explain_1, aes(x = prop, y = fct_rev(topic), fill = fct_rev(rating))) +
  geom_col(alpha = 0.8) +
  labs(x = "", y = "", fill = "",
       title = "The facilitator's explanation was easy to follow") +
  scale_fill_manual(values = c("#236EC3", "#78AFE6", "#E1E1E1", "#E6AAA0", "#C32314")) +
  guides(fill = guide_legend(reverse = TRUE)) + # reverse legend order
  scale_x_continuous(labels = scales::label_percent()) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank()
  )

p_materials <- ggplot(materials_1, aes(x = prop, y = fct_rev(topic), fill = fct_rev(rating))) +
  geom_col(alpha = 0.8) +
  labs(x = "", y = "", fill = "",
       title = "The training materials clear and adequate") +
  scale_fill_manual(values = c("#236EC3", "#78AFE6", "#E1E1E1", "#E6AAA0", "#C32314")) +
  guides(fill = guide_legend(reverse = TRUE)) + # reverse legend order
  scale_x_continuous(labels = scales::label_percent()) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank()
  ) 

```

```{r}
#| echo: false
#| message : false

# Analysis on feedback and respondents

# Load package

library(tidyverse)
library(googlesheets4)
library(stringr)
library(gtsummary)
library(flextable)

# Load data
url <- "https://docs.google.com/spreadsheets/d/12jxEV9rqIOrIw70lVXfSegorv25owciTMS-hjSotioo/edit?resourcekey#gid=1120526525"
dat <- read_sheet(url)
url2 <- "https://docs.google.com/spreadsheets/d/1ZMLHBtBn83r69u0IHmtAyjJ_QU8XV9uevYIuomj8x1g/edit#gid=146447740"

respondents <- dat |> 
  rename(email = 2) |> 
  mutate(use_ird_email = str_detect(email, "ird\\.global")) 

registrants <- read_sheet(url2) |> 
  rename(email = 2, 
         gender = 6,
         country = 7,
         type = 10,
         org = 11,
         SSRG = 16
         ) |> 
  filter(row_number() == n(), .by = email) |> # Take the last record if duplicated
  select(email, gender, country, type, org, SSRG) |> 
  mutate(email = recode(email, 'lapeonalaja@gmail.com' = 'lape.onalaja@ird.global', 'israr.ahmad.hri@nih.org.pk' = 'israrpmrc@gmail.com', 
  'Mondar.Ahmed@ird.global' = 'mondar.ahmed@ird.global', 'Maryam.younus@ird.global' = 'maryam.younus@ird.global',
  'aftab.khan.hri@nih.org.pk' = 'aftabropmrc@gmail.com', 'minhal.hyder@ird.golbal' = 'minhal.hyder@ird.global',
  'mohsin.h.alvi@pill.org.pk' = 'mohsinhassanalvi@gmail.com')) |> 
  mutate(country = recode(country, 'Karachi' = 'Pakistan'))

# merge details from registrant list to respondents 
combi <- respondents |> 
  left_join(registrants)

# flattened feedback of respondents and rename columns
feedbacks <- combi |> 
  group_by(email) |> 
  summarise_all(~paste(ifelse(is.na(.), "", .), collapse = "|")) |> 
  rename("Organisation - What was done well" = 7,
         "Organisation - Suggestions to improve" = 8,
         "Intro to R - What was done well" = 11,
         "Intro to R - Suggestions to improve" = 12,
         "Data Viz - What was done well" = 15,
         "Data Viz - Suggestions to improve" = 16,
         "Data Wrangling - What was done well" = 19,
         "Data Wrangling - Suggestions to improve" = 20,
         "Descriptive Analysis - What was done well" = 23,
         "Descriptive Analysis - Suggestions to improve" = 24,
         "Intro to Quarto - What was done well" = 27,
         "Intro to Quarto - Suggestions to improve" = 28,
         "Report generation - What was done well" = 31,
         "Report generation - Suggestions to improve" = 32,
         "General comments or suggestions" = 33
  ) |> 
  select(1,7,8,11,12,15,16,19,20,23,24,27,28,31,32,33)

# Summary table for registrants
registrants_details <- registrants |> 
  filter(row_number() == n(), .by = email) |> 
  select(2 : 6) |> 
  tbl_summary()
  
# Summary table for respondents
respondents_details <- combi |> 
  filter(row_number() == n(), .by = email) |>
  select(35 : 39) |> 
  tbl_summary()

#Splitting the feedback by sections
feedback_org <- feedbacks |> 
  select(starts_with("Organisation")) |> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |> 
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on pre-workshop Organisation")

feedback_intro <- feedbacks |> 
  select(starts_with("Intro to R")) |> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |> 
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Intro to R")

feedback_viz <- feedbacks |> 
  select(starts_with("Data Viz")) |> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |> 
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Data Viz")

feedback_wrang <- feedbacks |> 
  select(starts_with("Data Wrangling")) |> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |> 
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Data Wrangling")

feedback_descr <- feedbacks |> 
  select(starts_with("Descriptive")) |>
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |>
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Descriptive Analysis")

feedback_quarto <- feedbacks |> 
  select(starts_with("Intro to Quarto")) |> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |>
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Intro to Quarto")

feedback_report <- feedbacks |> 
  select(starts_with("Report generation"))|> 
  rename("What was done well" = 1,
         "Suggestions to improve" = 2) |>
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "Feedback on Report generation")

feedback_general <- feedbacks |> 
  select(starts_with("General")) |> 
  rename("Remarks" = 1) |> 
  flextable() |> 
  set_table_properties(layout = "autofit") |> 
  set_caption(caption = "General comments or suggestions")

# Summary table for registrants
registrants_details <- registrants |> 
  filter(row_number() == n(), .by = email) |> 
  select(2 : 6) |> 
  tbl_summary()
  
# Summary table for respondents
respondents_details <- combi |> 
  filter(row_number() == n(), .by = email) |>
  select(35 : 39) |> 
  mutate(
    gender = setLabels(gender, "Gender"),
    country = setLabels(country, "Country"),
    type = setLabels(type, "Type"),
    org = setLabels(org, "Organization"),
    SSRG = setLabels(SSRG, "Part of SSRG")
    )|> 
  tbl_summary() |> 
    bold_labels()
```

The Small Scale Research Grant is an initiative of the PIECEs project (funded by NIHR) under the mental health program-IRD-Pakistan. It aims to support early-career researchers in enhancing their research and grant management skills. Out of 111 received applications across Pakistan, five projects were finalized. The project teams received support through fortnightly check-ins from PIECEs team grant management officers. Additionally, knowledge exchange meetings with QMUL were organized to share updates and exchange practices with other SSRG-funded projects.

## Pre-registration:

The R session for SSRG was introduced as an extension of capacity-building training to SSRG awardees. The registration call for the R session was circulated in October 2023 among SSRG awardees and across IRD countries. Registration details included information on gender, country of residence, IRD employment status, SSRG awardee status, and consent for data privacy, future communication, and cancellation policy. A total of 48 participants from Bangladesh, the Philippines, Pakistan, Nigeria, the United States, the United Arab Emirates (Dubai), and South Africa registered for the training session.

## Workshop curriculum:

The training session took place successfully from January 22nd to 26th, spanning five days, with each session lasting 2.5 hours. The sessions were conducted by two esteemed facilitators, Zaw Myo Tun (Senior Manager, Global Research Office) and Cheng Yew Toon (Deputy Manager, Grant Management Office).

| **Topic**                                           | **Duration in hours** | **Presenter**  |
|----------------------------------|:-----------------:|:-----------------:|
| Intro to R syntax                                   |           1           | Cheng Yew Toon |
| Data viz (Including setting up project + read data) |          2.5          |  Zaw Myo Tun   |
| Data management                                     |          2.5          | Cheng Yew Toon |
| Factors and dates                                   |           1           |  Zaw Myo Tun   |
| Descriptive statistics                              |           1           |  Zaw Myo Tun   |
| Intro to Quarto                                     |           2           | Cheng Yew Toon |
| Programmatic report generation                      |           1           |  Zaw Myo Tun   |
| Basic dashboarding                                  |          1.5          |  Zaw Myo Tun   |

{{< pagebreak >}}

## Session attendance

The plot below shows daily attendance of R session participants.

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-align: center

dat <- data.frame(
  days = 1:5,
  attended = c(32, 31, 29, 25, 26)
)
```

```{r}
ggplot(dat, aes(x = days, y = attended)) +
  geom_line(colour = "steelblue") +
  geom_point(colour = "steelblue", size = 3) +
  labs(x = "workshop day", y = "",
       title = "Daily number of workshop participants") +
  scale_y_continuous(limits = c(0, 48), labels = scales::label_number()) +
  theme_minimal() +
  theme(
    panel.grid.minor.x = element_blank()
  )
```

In total, 35 people participated. 3 attended for one day, 1 attended for 2 days, and 31 attended for 3 or more days. `r respondents_count` participants provided feedback on what went well and areas for improvement. Certification was contingent on attending at least 3 out of the 5 days.

```{r}
r_session |> 
  select(gender, country, type, organization) |> 
  tbl_summary(label = list(
    gender ~ "Gender",
    country ~ "Country of Residence",
    type ~ "Type of Participant",
    organization ~ "Participant Organization"
  )) |> 
  bold_labels()
```

------------------------------------------------------------------------

## Feedback for R Workshop

Out of `r nrow(r_session)` participants who attended the workshop, `r respondents_count` provided their feedback for the workshop.

Below are the demographics of the feedback providers.

```{r}
#| echo: false
respondents_details
```

#### Feedback on Pre-workshop Organisation

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 2.5
#| fig-align: center
p_pre_wksp
```

Based on the feedback of the respondents, the pre-workshop organisation was seen as successful.

The participants were satisfied with the registration process, noting that the large majority of the respondents *strongly agreed* that it was easy to register for the workshop.

Large majority of the respondents *strongly agreed* that communications for the workshop were clear and they were timely.

Regarding the instructions for the workshop preparation, large majority of the respondents *strongly agreed* that it was clear.

The most common feedback on what was done well for the pre-workshop organisation is that the communication of the workshop was clear and timely.

From the data, many also gave general feedback about the whole workshop in this part of the feedback. It was encouraging to see that the respondents found the workshop to be well organised and informative. The training was found to be interactive and easy to follow.

Some areas of improvement that were suggested includes having more time to practice during the workshop and also have some breaks in between. `6` respondents mentioned that it would be good to have the workshop stretched out with assignments during the intervals, this is so that they can have time to practice.

------------------------------------------------------------------------

### Feedback on workshop activities and materials

In this section we will look at 2 likert scales for facilitator performance and quality of training materials. Subsequently, we will summarize the writtened feedback by the individual topics that were covered during the workshop.

#### Feedback on Facilitator performance

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 3
#| fig-align: center
p_explain
```

Regarding the facilitator's performance, the feedback are mostly well received.

-   For Intro to R, the large majority of the respondents *agreed* that the facilitator explanation was easy to follow.

-   For Data Visualization, the large majority of respondents *strongly agreed* that the facilitator explanation was easy to follow.

-   For Data Wrangling, the majority of respondents *strongly agreed* that the facilitator explanation was easy to follow. There is also a small proportion of participants who *disagreed*.

-   For Descriptive Analysis, the large majority of respondents *strongly agreed* that the facilitator explanation was easy to follow.

-   For Intro to Quarto, the large majority of respondents *strongly agreed* that the facilitator explanation was easy to follow.

-   For Report Generation, the large majority of respondents *strongly agreed* that the facilitator explanation was easy to follow.

#### Quality of training materials

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 3
#| fig-align: center
p_materials
```

Regarding the quality of training materials, the feedback are mostly well received as well.

-   For Intro to R, large majority of the respondents *agreed* that the training materials were clear and adequate.There is also a small proportion of participants who *disagreed*.

-   For Data Visualization, large majority of the respondents *strongly agreed* that the training materials were clear and adequate.

-   For Data Wrangling, large majority of the respondents *agreed* that the training materials were clear and adequate.

-   For Descriptive Analysis, large majority of the respondents *agreed* that the training materials were clear and adequate.

-   For Intro to Quarto, large majority of the respondents *strongly agreed* that the training materials were clear and adequate.

-   For Report Generation, large majority of the respondents *strongly agreed* that the training materials were clear and adequate.

------------------------------------------------------------------------

### Feedback by Topics

#### Intro to R

The most common feedback for the topic on Intro to R was that the contents were easy to follow and the session was well paced.

Other comments included were being the training materials were good and the session was information.

The most common suggestions for improvement was that the participants would like to have more hands-on practice and also have more time allocated to practice.

Other suggestions include having to share more materials for the participants to understand more in-depth and the session could be more in-depth.

------------------------------------------------------------------------

#### Data Visualization

The most common feedback for the topic on Data Visualization was that the topic was very useful and informative.

Other comments included were that the session was very hands-on and the session was well facilitated by having the participants share their screen and troubleshoot their roadblocks along the way.

It is worth noting that many participants highlighted that they enjoyed this session and they would like more time allocated to this topic with more hands-on practice.

Other areas of improvement mentioned was that it would be good to learn more about basic commands and syntax before going into data visualization.

------------------------------------------------------------------------

#### Data Wrangling

The most common feedback for the topic on Data Wrangling was that the topic was very useful and interactive.

Comments regarding areas of improvement included were that this topic is difficult, so it would be good to have time allocated for this topic and have more practice questions during the session.

------------------------------------------------------------------------

#### Descriptive Analysis

The most common feedback for the topic on Descriptive Analysis was that the topic was useful, the session was well delivered and well facilitated.

Most comments regarding areas of improvement were that the session was a little fast for them to follow.

Other suggested areas of improvement is to have more practice and allocate more time to this topic.

------------------------------------------------------------------------

#### Intro to Quarto

The most common feedback for the topic on Intro to Quarto was that the topic was useful.

Most comments regarding areas of improvement were that the participants would like more practice on this topic.

------------------------------------------------------------------------

#### Report Generation

The most common feedback for the topic on Report Generation was that the topic was interesting, useful and the session well demonstrated.

Most comments regarding areas of improvement were that the participants would like more time allocated to this topic and more practice opportunities during the session.

------------------------------------------------------------------------

#### General feedback and suggestions

The general feedback received was that the workshop was well planned and delivered. The participants were thankful for the learning opportunity and they would like more workshops like this.
